---
title: "fixed tree"
author: "Mathieu"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)
```

\newpage

# Datasets

Dataset         Sequences  Length  Coalescent          Clock   Substitution
------------   ---------- -------  ------------------- ------  ------------
ebov_dud17     1610         18992  Skygrid (50, 2.0)     UCLN  1+2+3(HKY+G4) + 14518:18992 HKY+G4
ebov_mba21      297         16757  Skygrid (50, 1.75)    UCLN  12+3(HKY+G4)  8 date unknown
fluH1L_bed15   2144          1695  Constant            Strict  12+3(HKY)
fluH3L_bed15   4006          1698  Constant            Strict  12+3(HKY)
fluPb2_wor14    354          2280  Skygrid (50, 150)    Local  12+3(HKY)+G4
fluVicL_bed15  1999          1755  Constant            Strict  12+3(HKY)
hiv_far14       927           438  Skygrid (50, 90.0)    UCLN  GTR+G4
lassaL_kli22    551          7038  Skygrid (50, 500.0)   UCLN  GTR+G4
mumps_mon21     467         15393  Skygrid (50, 25.0)  Strict  HKY+G4
rabies_via23    290         11883  Skygrid (50,20.0)     UCLN  12+3(GTR+G4)
sars2_can20    1046         29409  Skygrid (20, 0.5)   Strict  HKY+G4
sars2_lem21    3241         29409  Skygrid (20, 0.5)   Strict  HKY+G4
sars2_pek22     717         29232  Skygrid (20, 0.37)  Strict  GTR+I
wnv_del20       801         10302  Skygrid (50, 19.0)    UCLN  GTR+G4
zikv_gru19      283         10269  Skygrid (50, 5.0)     UCLN  1+2+3(HKY+G4)

\newpage

```{r}
source("../torchtree-testing/R/skygrid.R")
```

```{r}
ssstree_datasets <- "ssstree_datasets"
results <- file.path("results", "datasets")
```

```{r}
parse.dataset <- function(dataset, params, burnin) {
  param <- NULL
  sky <- NULL
  
  log.file <- file.path(ssstree_datasets, dataset, "param_run01.log")
  df <- read_tsv(log.file, skip = 3, show_col_types = FALSE)
  
  print(paste0("unconstrained: ", nrow(df)))
  df <- df[seq(1, nrow(df), 5), ]
  param <- df[as.integer(burnin * nrow(df)):nrow(df), ] %>%
    select(any_of(params)) %>%
    mutate(type = "unconstrained") %>%
    bind_rows(param)
  
  sky.file <- file.path(ssstree_datasets, dataset, "sky_run01.log")
  if (file.exists(sky.file)) {
    df <- read_tsv(sky.file, skip = 3, show_col_types = FALSE)
    
    sky <- df[as.integer(burnins[i] * nrow(df)):nrow(df), ] %>%
      select(starts_with("skygrid.logPopSize")) %>%
      mutate(type = "unconstrained") %>%
      bind_rows(sky)
  }
  
  for (topology in c("unconstrained", "lsd2", "treetime", "map", "mcc", "rerooted")) {
    stem = file.path("results", "datasets", dataset, topology)
    log.file = file.path(stem, "param_run01.log")
    
    df.log <- read_tsv(log.file, skip = 3, show_col_types = FALSE)
    
    print(paste0(stem, " ", topology, ": ", nrow(df.log)))
    param2 <- df.log %>%
      slice(seq(floor(0.5 * n()), n(), by = 5)) %>%
      select(any_of(params)) %>%
      mutate(type = topology)
    
    print(paste0(topology, ": ", nrow(param2)))
    
    if (file.exists(file.path(stem, "sky_run01.log"))) {
      df <- read_tsv(
        file.path(stem, "sky_run01.log"),
        skip = 3,
        show_col_types = FALSE
      )
      
      sky <- df %>%
        select(starts_with("skygrid.logPopSize")) %>%
        mutate(type = topology) %>%
        slice(seq(floor(0.5 * n()), n(), by = 5)) %>%
        bind_rows(sky)
    }
    param <- rbind(param, param2)
  }
  
  lsd = readLines(file.path(results, dataset, "lsd2", "ali.fasta.lsd"))
  lsd = lsd[grepl(" rate ", lsd) & grepl(" tMRCA ", lsd)]
  
  treetime = readLines(
    file.path(
      "results-treetime",
      "datasets",
      dataset,
      "treetime",
      "treetime",
      "molecular_clock.txt"
    )
  )
  treetime = strsplit(treetime[grepl("rate", treetime)], "\t")[[1]][2]
  
  save(param, sky, lsd, treetime, file = paste0(dataset, ".RData"))
  
  list(
    param = param,
    sky = sky,
    lsd = lsd,
    treetime = treetime
  )
}

prepare.skygrids <- function(param, sky, cutoff, min, max, youngest) {
  skygrids <- NULL
  
  param.split <- param %>%
    group_by(type) %>%
    group_split()
  
  sky.split <- sky %>%
    group_by(type) %>%
    group_split()
  
  for (i in 1:length(param.split)) {
    root_height <- param.split[[i]] %>% select(rootHeight)
    log_pop_size <- sky.split[[i]] %>% select(starts_with("skygrid.logPopSize"))
    skygrid <- prepareSkygrid(root_height, log_pop_size, cutoff, min, max, youngest)
    # skygrid <- prepare_skygrid(root_height, log_pop_size, cutoff, age_of_youngest=youngest)
    skygrids <- rbind(skygrids, mutate(skygrid, type = param.split[[i]]$type[1]))
  }
  skygrids %>% mutate(
    low = exp(low),
    high = exp(high),
    median = exp(median)
  )
  # skygrid <-  mutate(skygrid, across(all_of(c("trajectory", "trajectory_low", "trajectory_high"))), exp)
}
```

\newpage

# ebov_dud17

```{r ebov_dud17-parse}
params <- c("rootHeight", "meanRate", "treeLength", "cp1.kappa", "cp1.alpha", "cp1.mu", "skygrid.precision")
data <- parse.dataset("ebov_dud17", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 0.5, 2019.5, 2020.4, 2020.327868852459)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# ebov_mba21

```{r ebov_mba21-parse}
params <- c("rootHeight", "meanRate", "treeLength", "cp12.alpha", "cp12.mu", "skygrid.precision")
data <- parse.dataset("ebov_mba21", params, 0.5)
```
```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 1.75, 2018, 2020.2, 2020.1284153005465)
```

```{r}
# skyplot(skygrids, fill_by=type, color = type, alpha=0.0) + ylab("Effective population size")
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 1.75, 1.75/49)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# fluH1L_bed15

```{r fluH1_bed15-parse}
params <- c("rootHeight", "clock.rate", "treeLength", "constant.popSize", "cp12.mu", "cp12.kappa")
data <- parse.dataset("fluH1L_bed15", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

\newpage

# fluH3L_bed15

```{r fluH3L_bed15-parse}
params <- c("rootHeight", "clock.rate", "treeLength", "constant.popSize", "cp12.kappa", "cp12.mu")
data <- parse.dataset("fluH3L_bed15", params, 0.5)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

\newpage

# fluPb2_wor14

```{r fluPb2_wor14-parse}
params <- c(
  "rootHeight", "treeLength", "cp12.alpha", "cp12.kappa", "cp12.mu",
  "tmrca(equine1)", "tmrca(equine2)", "tmrca(galliform1)", "tmrca(human1)", "tmrca(shorebird1)", "tmrca(swine1)", "tmrca(swine2)", "tmrca(swine3)",
  "avian.rate", "equine.rate", "galliform1.rate", "human1.rate", "shorebird1.rate", "swine1.rate", "swine2.rate", "swine3.rate", "skygrid.precision"
)
data <- parse.dataset("fluPb2_wor14", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 150, 1850.5, 2011.4, 2011)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# fluVicL_bed15

```{r fluVicL_bed15-parse}
params <- c("rootHeight", "treeLength", "clock.rate", "constant.popSize", "cp12.kappa", "cp12.mu")
data <- parse.dataset("fluVicL_bed15", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

\newpage

# hiv_far14

```{r hiv_far14-parse}
params <- c("rootHeight", "treeLength", "meanRate", "alpha", "gtr.rates.rateAC", "frequencies1", "skygrid.precision")
data <- parse.dataset("hiv_far14", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 90, 1915, 2005.5, 2005)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# lassaL_kli22

```{r lassaL_kli22-parse}
params <- c("rootHeight", "meanRate", "treeLength", "alpha", "gtr.rates.rateAC", "skygrid.precision")
data <- parse.dataset("lassaL_kli22", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 1519.0, 1019, 2019.2, 2019.1780821917807)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# mumps_mon21

```{r mumps_mon21-parse}
params <- c("rootHeight", "treeLength", "clock.rate", "alpha", "kappa", "frequencies1", "skygrid.precision")
data <- parse.dataset("mumps_mon21", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 25, 1993, 2018.5, 2018.0)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# rabies_via23

```{r rabies_via23-parse}
params <- c("rootHeight", "meanRate", "treeLength", "cp12.alpha", "cp12.mu", "skygrid.precision")
data <- parse.dataset("rabies_via23", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 20.0, 1998.75, 2016.1, 2016.0738)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# sars2_can20

```{r sars2_can20-parse}
params <- c("rootHeight", "clock.rate", "treeLength", "alpha", "kappa", "skygrid.precision")
data <- parse.dataset("sars2_can20", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 0.5, 2019.5, 2020.4, 2020.327868852459)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# sars2_lem21

```{r sars2_lem21-parse}
params <- c("rootHeight", "treeLength", "clock.rate", "alpha", "kappa", "frequencies1", "skygrid.precision")
data <- parse.dataset("sars2_lem21", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 0.8, 2020, 2020.9, 2020.827868852459)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# sars2_pek22

```{r sars2_pek22-parse}
params <- c("rootHeight", "clock.rate", "treeLength", "pInv", "skygrid.precision")
data <- parse.dataset("sars2_pek22", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 0.37, 2019.75, 2020.2, 2020.120218579235)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# wnv_del20

```{r wnv_del20-parse}
params <- c("rootHeight", "treeLength", "meanRate", "alpha", "gtr.rates.rateAC", "frequencies1", "skygrid.precision")
data <- parse.dataset("wnv_del20", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 19, 1998, 2017, 2016.6475)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```

\newpage

# zikv_gru19

```{r zikv_gru19-parse}
params <- c("rootHeight", "meanRate", "treeLength", "meanRate", "cp1.alpha", "cp1.mu", "skygrid.precision")
data <- parse.dataset("zikv_gru19", params, 0.5)
```

```{r, results='asis'}
cat(data$lsd)
```

```{r}
df <- data$param %>% gather("param", "value", -type)
ggplot(df, aes(value, group = type, colour = type)) +
  geom_density() +
  facet_wrap(~param, scales = "free")
```

```{r}
skygrids <- prepare.skygrids(data$param, data$sky, 5.0, 2013, 2018.1, 2018.0465753424658)
```

```{r}
ggplot(skygrids, aes(x = x, y = median, color = type)) +
  theme_bw() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(), strip.background = element_blank(),
    panel.border = element_rect(colour = "black")
  ) +
  geom_ribbon(aes(ymin = low, ymax = high, fill = type), alpha = 0.0) +
  geom_line() +
  # geom_vline(xintercept=2020.1284153005465-seq(0, 0.37, 0.37/19)) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x))
  ) +
  xlab("Year") +
  ylab("Effective population size") +
  labs(title = "skygrids")
```
