---
title: "fixed tree: summarry"
author: "Mathieu"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(conflicted)
library(ggpubr)
library(scales)
library(ggbreak)

conflict_prefer("filter", "dplyr")
#library(scales)
knitr::opts_chunk$set(fig.path='Figs/')
```

\newpage

```{r}
kld_base = function(x, y, ...) {
  integrand = function(x, y, t) {
    f.x =  approx(density(x)$x, density(x)$y, t)$y
    f.y =  approx(density(y)$x, density(y)$y, t)$y
    tmpRatio = f.x * (log2(f.x) - log2(f.y))
    tmpRatio = ifelse(is.infinite(tmpRatio), 0, ifelse(is.na(tmpRatio), 0, tmpRatio))
    return(tmpRatio)
  }
  return(integrate(
    integrand,
    0,
    Inf,
    x = x,
    y = y,
    stop.on.error = FALSE
  )$value)
}

kl_l <- function(x, y) {
  f.x <- approxfun(density(x, bw = "nrd0"))
  f.y <- approxfun(density(y, bw = "nrd0"))
  integrand <- function(z) {
    return((log(f.x(z)) - log(f.y(z))) * f.x(z))
  }
  return(integrate(integrand, lower = 2, upper = 100))
}
```


```{r eval=FALSE, include=FALSE}
datasets = c("fluH3L_bed15", "lassaL_kli22", "wnv_del20", "sars2_lem21", "ebov_dud17",
  "ebov_mba21", "fluH1L_bed15", "fluPb2_wor14", "fluVicL_bed15", "hiv_far14",
  "mumps_mon21", "rabies_via23", "sars2_can20", "sars2_pek22", "zikv_gru19")
all = NULL
all_mean = NULL
all_sky = NULL
for (i in seq_along(datasets)) {
  dataset = datasets[[i]]
  load(paste0(dataset, ".RData"))
  rate_name <- intersect(c("meanRate", "clock.rate", "avian.rate"), names(param))
  param = param %>% mutate(!!"rate" := .[[rate_name[1]]]) %>% filter(type !="fixed")
  
  df = param %>% group_by(type) %>%
    summarize(
      rootHeightMean = mean(rootHeight),
      rootHeightVar = var(rootHeight),
      rateMean = mean(rate)
    ) %>%
    mutate(dataset = dataset)

  all_mean = param %>% group_by(type) %>%
    summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
    pivot_longer(!type, names_to = "parameter", values_to = "value") %>%
    mutate(dataset = dataset) %>%
    bind_rows(all_mean)
  
  all = rbind(all, df)
  #assign(paste0("param_", dataset), get("param"))
  #assign(paste0("sky_", dataset), get("sky"))
  #assign(paste0("lsd_", dataset), get("lsd"))
  if (!is.null(sky)) {
    prefixes <- names(sky)[1:length(sky) - 1]
    m = sky %>% group_by(type) %>%
      summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)))
    uncons = m %>% filter(type == "unconstrained1") %>%
      pivot_longer(cols = all_of(prefixes),
                   names_to = "pop",
                   values_to = "value")
    cons = m %>% filter(type != "unconstrained1") %>%
      pivot_longer(cols = all_of(prefixes),
                   names_to = "pop",
                   values_to = "value")
    
    df = cons %>% full_join(uncons, by = "pop") %>%
      mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
      select(type.x, value) %>%
      # RRMSE relative root mean squared error
      group_by(type.x) %>% summarize(across(where(is.numeric), ~ sqrt(mean(.)))) %>%
      rename(Method = type.x, RRMSE = value) %>%
      select(Method, RRMSE) %>%
      mutate(Dataset = dataset)
    
    all_sky = df %>% bind_rows(all_sky)
  }
}
```

```{r eval=FALSE, include=FALSE}
datasets = c(
  "fluH3L_bed15",
  "lassaL_kli22",
  "wnv_del20",
  "sars2_lem21",
  "ebov_dud17",
  "ebov_mba21",
  "fluH1L_bed15",
  "fluPb2_wor14",
  "fluVicL_bed15",
  "hiv_far14",
  "mumps_mon21",
  "rabies_via23",
  "sars2_can20",
  "sars2_pek22",
  "zikv_gru19"
)

all_mean = NULL
all_sd = NULL
sky_mean = NULL
sky_sd = NULL

for (dataset in datasets) {
  for (topology in c("unconstrained", "lsd2", "treetime", "map", "mcc", "rerooted")) {
    if (topology == "unconstrained") {
      stem = file.path("ssstree_datasets", dataset)
      log.file = file.path(stem, paste0("param_run0", 1, ".log"))
    } else{
      stem = file.path("results", "datasets", dataset, topology)
      log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      if (!file.exists(log.file)) {
        stem = file.path("results-lsd", "datasets", dataset, topology)
        log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      }
      if (!file.exists(log.file)) {
        stem = file.path("results-treetime", "datasets", dataset, topology)
        log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      }
      if (!file.exists(log.file)) {
        stem = file.path("results-treetime", "datasets", dataset)
        log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      }
    }
    if (!file.exists(log.file)) {
      next
    }
    
    df.log <- read_tsv(log.file, skip = 3, show_col_types = FALSE)
    df.log = df.log %>%
      slice(seq(floor(0.5 * n()), n(), by=5)) %>%
      select(-state)
    #df.log <- df.log[seq(1, nrow(df.log), 5), ] # subsample
    #df.log <- df.log[as.integer(0.5 * nrow(df.log)):nrow(df.log), ] %>% select(-state)
    
    all_mean = df.log %>%
      summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
      pivot_longer(cols = everything(),
                   names_to = "parameter",
                   values_to = "value") %>%
      mutate(dataset = dataset, type = topology) %>%
      bind_rows(all_mean)
    
    all_sd = df.log %>%
      summarise(across(where(is.numeric), ~ sd(.x, na.rm = TRUE))) %>%
      pivot_longer(cols = everything(),
                   names_to = "parameter",
                   values_to = "value") %>%
      mutate(dataset = dataset, type = topology) %>%
      bind_rows(all_sd)
    
    if (file.exists(file.path(stem, paste0("sky_run", "0", 1, ".log")))) {
      sky.log <- read_tsv(file.path(stem, paste0("sky_run", "0", 1, ".log")),
                          skip = 3,
                          show_col_types = FALSE)
      sky.log = sky.log %>%
        slice(seq(floor(0.5 * n()), n(), by=5)) %>%
        select(-state)

      sky_mean = sky.log %>%
        summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) %>%
        pivot_longer(
          cols = everything(),
          names_to = "pop",
          values_to = "value"
        ) %>%
        mutate(dataset = dataset, type = topology) %>%
        bind_rows(sky_mean)
      
      sky_sd = sky.log %>%
        summarise(across(where(is.numeric), \(x) sd(x, na.rm = TRUE))) %>%
        pivot_longer(
          cols = everything(),
          names_to = "pop",
          values_to = "value"
        ) %>%
        mutate(dataset = dataset, type = topology) %>%
        bind_rows(sky_sd)
    }
  }
}
save(all_mean, all_sd, sky_mean, sky_sd, file = "summary.RData")
```

```{r}
load("summary.RData")
all_mean$type <- recode(all_mean$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="MRC")

all_sd$type <- recode(all_sd$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="MRC")

sky_mean$type <- recode(sky_mean$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="MRC")
```

```{r}
all_mean = all_mean %>% mutate(stat="mean")
all_sd = all_sd %>% mutate(stat="sd")

uncons = all_mean %>% bind_rows(all_sd) %>% filter(type == "unconstrained") %>% select(-type)
cons = all_mean %>% bind_rows(all_sd) %>% filter(type != "unconstrained")

df_all = cons %>% full_join(uncons, by = c("dataset", "parameter", "stat")) %>%
  mutate(value = (value.x - value.y) / value.y) %>%
  select(type, parameter, dataset, value, stat) %>%
  rename(Method = type, Dataset = dataset) %>%
  filter(parameter != "rate") %>%
  mutate(
    model = case_when(
      parameter == "meanRate" |
        parameter == "clock.rate" | str_detect(parameter, ".rate$") ~ "clock",
      parameter == "rootHeight" |
        parameter == "treeLength" | str_detect(parameter, "^tmrca") ~ "tree",
      parameter == "constant.popSize" |
        parameter == "skygrid.precision" ~ "coalescent",
      str_detect(parameter, "gtr") |
        str_detect(parameter, "mu") |
        str_detect(parameter, "frequencies") |
        str_detect(parameter, "kappa") ~ "subsmodel",
      str_detect(parameter, "mu") |
        str_detect(parameter, "alpha") | parameter == "pInv" ~ "sitemodel",
      TRUE                            ~ "other"
    )
  )
df_mean = df_all %>% filter(stat == "mean")
df_sd = df_all %>% filter(stat == "sd")
```

```{r}
rerooted = df_mean %>% filter(Method == "MRC")
same = df_mean %>% filter(Method != "MRC") %>% filter(Method %in% c("LSD", "TreeTime"))

df2=rerooted %>% left_join(same, by = c("Dataset", "parameter", "stat", "model")) %>%
  rename(rerooted=value.x, heuristic=value.y) %>%
  mutate(diff=rerooted-heuristic) %>%
  filter(model!= "other")

ggplot(df2, aes(rerooted, heuristic, col=model, shape=Method.y))+
  geom_point() +
  geom_abline(slope=1,intercept=0) #+
  #facet_wrap(~model)
```
```{r}
print(paste0(nrow(filter(df2, Method.y=="LSD" & diff < 0)), " ", nrow(filter(df2, Method.y=="LSD"))))
print(paste0(nrow(filter(df2, Method.y=="TreeTime" & diff < 0)), " ", nrow(filter(df2, Method.y=="TreeTime"))))

print(paste0(nrow(filter(df2, Method.y=="LSD" & diff < 0 & model %in% c("coalescent", "tree", "clock"))), " ", nrow(filter(df2, Method.y=="LSD",model %in% c("coalescent", "tree", "clock")))))
print(paste0(nrow(filter(df2, Method.y=="TreeTime" & diff < 0 & model %in% c("coalescent", "tree", "clock"))), " ", nrow(filter(df2, Method.y=="TreeTime",model %in% c("coalescent", "tree", "clock")))))
```

```{r}
df2=cons %>% full_join(uncons, by = c("dataset", "parameter", "stat")) %>%
  rename(constrained=value.x, unconstrained=value.y) %>%
  filter(parameter!="state"& parameter!="treeLength") %>%
  mutate(
    model = case_when(
      parameter == "meanRate" |
        parameter == "clock.rate" | str_detect(parameter, ".rate$") ~ "clock",
      parameter == "rootHeight" |
        parameter == "treeLength" | str_detect(parameter, "^tmrca") ~ "tree",
      parameter == "constant.popSize" |
        parameter == "skygrid.precision" ~ "coalescent",
      str_detect(parameter, "gtr") |
        str_detect(parameter, "mu") |
        str_detect(parameter, "frequencies") |
        str_detect(parameter, "kappa") ~ "subsmodel",
      str_detect(parameter, "mu") |
        str_detect(parameter, "alpha") | parameter == "pInv" ~ "sitemodel",
      TRUE                            ~ "other"
    )
  ) %>%
  filter(stat=="mean")

ggplot(df2, aes(x=unconstrained, y=constrained, col=type, shape=model)) +
  geom_point() +
  geom_abline(slope=1,intercept=0)
```

# Substitution and site model parameters

- Substitution model: kappa, gtr.rates, frequencies
- Site model: alpha, pInv, mu

```{r insensitive}
df2 = df_mean %>% filter(model %in% c("subsmodel", "sitemodel"))

df2$model = recode_factor(df2$model, "sitemodel"="Site", "subsmodel"="Substitution", "clock"="Clock", "tree"="Tree", "coalescent"="Coalescent")
df2$Method = factor(df2$Method, levels=c("MCC", "MSP", "MRC", "LSD", "TreeTime"))

ggplot(df2, aes(y = value, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "Relative error")
```

```{r}
df2 %>%
  mutate(group = case_when(
    value < -0.25             ~ "< -0.25",
    value <= 0.25             ~ "[-0.25, 0.25]",
    value > 0.25              ~ "> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))
```

# Clock, coalescent, tree parameters/statistics

- Coalescent: constant popSize, precision of GMRF
- Tree: rootHeight, treeLength, tmrcas
- Clock model: strict clock rate, meanRate (UCLN), *.rate of local clock

```{r}
df2 = df_mean %>% filter(model %in% c("coalescent", "tree", "clock"))  %>%
  #filter(!parameter %in% c("meanRate", "treeLength"))
  filter(!parameter %in% c("treeLength"))
df2$Method = factor(df2$Method, levels=c("MCC", "MSP", "MRC", "LSD", "TreeTime"))
df2$model <- recode(df2$model,
                        "clock" = "Clock",
                        "coalescent" = "Coalescent",
                        "tree" = "Tree")

ggplot(df2, aes(y = value, col = Method, x = Dataset)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  geom_point(aes(shape = model, group = model), position = position_dodge(width = 0.6)) +
  theme_bw() +
  #scale_y_log10() +
  #scale_y_break(c(7, 13)) +scale_y_break(c(15,32)) +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    #strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "Relative error")
```

```{r sensitive}
df2$value2 = sign(df2$value) * log(1+abs(df2$value))
signed_log <- function(y) sign(y) * log(1 + abs(y))

ggplot(df2, aes(y = value2, col = Method, x = Dataset)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  geom_point(aes(shape = model, group = model), position = position_dodge(width = 0.6)) +
  geom_hline(yintercept=signed_log(0.25), linetype = "dashed") +
  geom_hline(yintercept=signed_log(-0.25),linetype = "dashed") +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    #strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = expression("Relative error (signed-log" [e] * " scale)")) +
  scale_y_continuous(
    breaks = signed_log(c(-5, -1, 0, 1, 5, 10, 20, 30)),  # where to place them
    labels = c(-5, -1, 0, 2, 5, 10, 20, 30)               # what to show
  )
```

```{r}
df2 %>%
  mutate(group = case_when(
    value < -0.25             ~ "< -0.25",
    value <= 0.25             ~ "[-0.25, 0.25]",
    value > 0.25              ~ "> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))

df2 %>% filter(Method %in% c('LSD','TreeTime')) %>%
  mutate(group = case_when(
    value < -0.25             ~ "< -0.25",
    value <= 0.25             ~ "[-0.25, 0.25]",
    value > 0.25              ~ "> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))

df2 %>% filter(Method %in% c('MSP')) %>%
  mutate(group = case_when(
    value < -0.25             ~ "< -0.25",
    value <= 0.25             ~ "[-0.25, 0.25]",
    value > 0.25              ~ "> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))

df2 %>% filter(Method %in% c('MSP','MCC')) %>%
  mutate(group = case_when(
    value < -0.25             ~ "< -0.25",
    value <= 0.25             ~ "[-0.25, 0.25]",
    value > 0.25              ~ "> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))

df2 %>% select(-stat, -value2) %>%
  pivot_wider(
    id_cols = c(parameter, Dataset, model),
    names_from = Method,
    values_from = value
  ) %>%
  filter(!is.na(MRC))
```

```{r re-mean}
df2 = df_mean %>%
  filter(model %in% c("coalescent", "tree", "clock", "subsmodel", "sitemodel")) %>%
  filter(!parameter %in% c("meanRate", "treeLength"))

df2$model = recode_factor(
  df2$model,
  "sitemodel" = "Site",
  "subsmodel" = "Substitution",
  "clock" = "Clock",
  "tree" = "Tree",
  "coalescent" = "Coalescent"
)
df2$Method = factor(df2$Method, levels = c("MCC", "MSP", "MRC", "LSD", "TreeTime"))

ggplot(df2, aes(y = value, x = model, col = Method)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = Dataset)) +
  geom_boxplot() +
  #geom_jitter(size = 0.5) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.1, 0.8),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank()
  ) +
  labs(x="Model", y = "Relative error") #+ ggtitle("Relative error of mean")
```

```{r}
df2 = df_sd %>% filter(model %in% c("coalescent", "tree", "clock", "subsmodel", "sitemodel")) %>%
  filter(!parameter %in% c("meanRate", "treeLength"))

df2$model = recode_factor(df2$model, "sitemodel"="Site", "subsmodel"="Substitution", "clock"="Clock", "tree"="Tree", "coalescent"="Coalescent")
df2$Method = factor(df2$Method, levels=c("MCC", "MSP", "MRC", "LSD", "TreeTime"))

ggplot(df2, aes(y = value, col = Method, x = model)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = Dataset)) +
  geom_boxplot() +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "Relative error") +
  ggtitle("Relative error of standard deviation")
```

# Skygrid population size parameters

$RRMSE(\bf{\bar x}) = \sqrt{1/n \sum_i^n \left(\frac{\bar x_i - \bar x_i^{u}}{\bar x_i^{u}}\right)^2}$ where $\bar x_i^{u}$ is the $i^{th}$ mean population size parameter of the unconstrained run and $\bar x_i$ is the mean population size parameter for a constrained run.

```{r skygrids}
uncons = sky_mean %>% filter(type == "unconstrained")
cons = sky_mean %>% filter(type != "unconstrained")

df.pop = cons %>% full_join(uncons, by = c("dataset", "pop")) %>%
  mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
  select(type.x, value, dataset) %>%
  group_by(type.x, dataset) %>% summarise(sqrt_mean = sqrt(mean(value)), .groups = "drop") %>%
  rename(Method = type.x, RMSRE = sqrt_mean, Dataset = dataset)# %>% filter(RRMSE <100)


ggplot(df.pop, aes(y = RMSRE, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1) +
  geom_hline(yintercept=0.25,linetype = "dashed") +
  scale_y_log10(breaks = c(0.001,0.01, 0.1, 1, 10, 100),
  labels = c("0.001", "0.01", "0.1", "1", "10", "100")) +
  labs(y = expression("RMSRE (log" [10] * " scale)")) +

  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

```{r}
res=df.pop %>%
  mutate(MethodGroup = case_when(
    Method %in% c("LSD", "TreeTime") ~ "LSDtreetime",
    Method %in% c("MCC", "MSP")      ~ "MCCMSP",
    TRUE                             ~ Method
  )) %>%
  #group_by(MethodGroup, Dataset) %>%
group_by(MethodGroup) %>%
  summarise(mean_RMSE = mean(RMSRE), .groups = "drop")
res
res[[2]][1] /res[[2]][2]
```

```{r}
df.pop %>%
  mutate(group = case_when(
    RMSRE < 0.25             ~ "< 0.25",
    RMSRE >= 0.25              ~ "=> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))

df.pop %>% filter(Method %in% c('LSD','TreeTime')) %>%
  mutate(group = case_when(
    RMSRE < 0.25             ~ "< 0.25",
    RMSRE >= 0.25              ~ "=> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))

df.pop %>% filter(Method %in% c('MSP','MCC')) %>%
  mutate(group = case_when(
    RMSRE < 0.25             ~ "< 0.25",
    RMSRE >= 0.25              ~ "=> 0.25"
  )) %>%
  count(group)%>%
  mutate(percent = 100 * n / sum(n))

df.pop %>%
  pivot_wider(
    #id_cols = c(parameter, Dataset, model),
    names_from = Method,
    values_from = RMSRE
  ) %>%
  filter(!is.na(MRC))
```

```{r}
df3 = df_mean %>% mutate(value=value**2)

uncons = sky_mean %>% filter(type == "unconstrained")
cons = sky_mean %>% filter(type != "unconstrained")

df.pop = cons %>% full_join(uncons, by = c("dataset", "pop")) %>%
  mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
  select(type.x, value, dataset) %>%
  group_by(type.x, dataset) %>% summarise(sqrt_mean = sqrt(mean(value)), .groups = "drop") %>%
  rename(Method = type.x, RRMSE = sqrt_mean, Dataset = dataset) %>%
  mutate(model="coalescent2", stat="mean", parameter="skygrid") %>% rename(value=RRMSE)
  
df2 = df.pop %>%
  bind_rows(df3) %>% filter(model %in% c("coalescent", "coalescent2", "tree", "clock"))  %>%
  filter(!parameter %in% c("meanRate", "treeLength"))
  #filter(!parameter %in% c("treeLength"))

df2$Method = factor(df2$Method, levels=c("MCC", "MSP", "MRC", "LSD", "TreeTime"))

#pdf("Figs/sensitive.pdf")
ggplot(df2, aes(y = value, col = Method, x = Dataset)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  geom_point(aes(shape = model, group = model), position = position_dodge(width = 0.6)) +
  theme_bw() +
  scale_y_log10() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "Relative error")
```


```{r}
my_colors <- c(
  "MCC" = "#0072B2",  # Blue
  "MSP" = "#E69F00",  # Orange
  "MRC" = "#009E73",  # Green
  "LSD" = "#D55E00",  # Red
  "TreeTime" = "#CC79A7",  # Purple
  "Unconstrained" = "#999999"   # Grey
)
my_breaks = c("MCC", "MSP", "MRC", "LSD", "TreeTime", "Unconstrained")

df_params = df_mean %>% filter(!parameter %in% c("treeLength")) %>% mutate(value=value**2) %>%
  group_by(model,Dataset,Method)%>% summarize(value=mean(value), .groups = "drop") %>%
  mutate(stat="mean")

uncons = sky_mean %>% filter(type == "unconstrained")
cons = sky_mean %>% filter(type != "unconstrained")

df.pop = cons %>% full_join(uncons, by = c("dataset", "pop")) %>%
  mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
  select(type.x, value, dataset) %>%
  group_by(type.x, dataset) %>% summarise(sqrt_mean = sqrt(mean(value)), .groups = "drop") %>%
  rename(Method = type.x, RRMSE = sqrt_mean, Dataset = dataset) %>%
  mutate(model="coalescent", stat="mean") %>% rename(value=RRMSE)

df2 = df.pop %>%
  bind_rows(df_params) %>%
  filter(model %in% c("coalescent", "tree", "clock"))

df2$Method = factor(df2$Method,
                    levels = c("Unconstrained", "MCC", "MSP", "MRC", "LSD", "TreeTime"))
df2$model = recode_factor(
  df2$model,
  "sitemodel" = "Site",
  "subsmodel" = "Substitution",
  "clock" = "Clock",
  "tree" = "Tree",
  "coalescent" = "Coalescent"
)

order_dataset = df2 %>% group_by(Dataset) %>% summarize(mean_value = mean(value)) %>% arrange(desc(mean_value))
#order_dataset <- rbind(head(order_dataset, 4), tail(order_dataset, 4))
order_dataset <- rbind(head(order_dataset, 2), tail(order_dataset, 2))
df2 = df2 %>% filter(Dataset %in% order_dataset$Dataset)
                     
#df2 = df2 %>% filter(Dataset %in% order_dataset[1:4] & Dataset %in% order_dataset[12:15])

df2$Dataset = factor(df2$Dataset, order_dataset$Dataset)


highlight_points <- df.pop %>% filter(Dataset=='ebov_mba21', model=="coalescent")

highlight <- data.frame(
  #xmin = min(highlight_points$x),
  #xmax = max(highlight_points$x),
  xmin = 1 - 0.4,
  xmax = 1 + 0.4,
  ymin = min(highlight_points$value),
  ymax = max(highlight_points$value)
)

highlight_points_fluPb2_wor14 <- df.pop %>% filter(Dataset=='fluPb2_wor14', model=="coalescent")

highlight_fluPb2_wor14 <- data.frame(
  #xmin = min(highlight_points$x),
  #xmax = max(highlight_points$x),
  xmin = which(order_dataset=='fluPb2_wor14') - 0.4,
  xmax = which(order_dataset=='fluPb2_wor14') + 0.4,
  ymin = min(highlight_points_fluPb2_wor14$value),
  ymax = max(highlight_points_fluPb2_wor14$value)
)

yLabeler <- function(x) {
    sapply(x, function(val) {
      if (val >= 1) {
        scales::comma(val, accuracy = 1)
      } else if (val >= 0.01) {
        scales::comma(val, accuracy = 0.01)
      } else if (val >= 0.001) {
        scales::comma(val, accuracy = 0.001)
      } else if (val >= 0.0001) {
        scales::comma(val, accuracy = 0.0001)
      } else if (val >= 1e-5) {
        scales::comma(val, accuracy = 1e-5)
      } else {
        format(val, scientific = TRUE)
      }
    })
}

yLabeler <- function(x) {
  sapply(x, function(xi) {
    log10xi <- log10(xi)
    if (abs(log10xi - round(log10xi)) < 1e-8) {
      # exact power of ten
      parse(text = paste0("10^", round(log10xi)))
    } else {
      # otherwise, just show number
      format(xi, scientific = TRUE)
    }
  })
}

scatter_plot = ggplot(df2, aes(x = Dataset, y = value, color = Method)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  geom_point(aes(shape = model, group = model), position = position_dodge(width = 0.5)) +
  # geom_point(
  #    data = data.frame(value = NA, Method = "Unconstrained"),
  #    aes(x = value, y = NA, color = Method),
  #    inherit.aes = FALSE
  #  ) +
  geom_point(
    data = data.frame(Dataset = df2$Dataset[1], value = df2$value[1], Method = "Unconstrained", model = df2$model[1]),
    aes(x = Dataset, y = value, color = Method)
  ) +
  theme_bw() +
  #scale_y_log10(breaks=c(1000,10, 0.1, 0.001,1e-5), labels=yLabeler) +
  scale_y_log10(breaks = c(1000,10, 0.1, 0.001,1e-5), labels =yLabeler)+
  #scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = math_format(10^.x))+
  scale_color_manual(values = my_colors, breaks=my_breaks, guide = guide_legend(title = "Method", nrow = 1)) +
  scale_shape_manual(
    values = c("Clock" = 16, "Tree" = 17, "Coalescent" = 15),  # Example shapes
    guide = guide_legend(title = "Model", nrow = 1)
  ) +
  # guides(
  #   color = guide_legend(order = 1),
  #   shape = guide_legend(order = 2)
  # ) +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.position="top",
    legend.box = "vertical",           # Stack legends vertically
    legend.spacing.y = unit(-0.5, "cm"),
    legend.box.just = "left",
    legend.spacing.x = unit(1, "cm"),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
    # geom_rect(data = highlight, inherit.aes = FALSE,
    #         aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    #         fill = NA, color = "red", linewidth = 1) +
    # geom_rect(data = highlight_fluPb2_wor14, inherit.aes = FALSE,
    #         aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    #         fill = NA, color = "red", linewidth = 1) +
  labs(y = "Mean squared relative error")
  #scale_y_continuous(labels = scales::comma) +
#annotate("text", x = 10, y = 1.5, label = "Legend for Methods and Models", size = 6, hjust = 0)
scatter_plot
```

```{r eval=FALSE, include=FALSE}
# Step 1: Store the true x-axis values
dataset_levels <- unique(df2$Dataset)

# Step 2: Build plot
scatter_plot <- ggplot(df2) +
  geom_point(
    aes(shape = model, group = model, x = Dataset, y = value, color = Method),
    position = position_dodge(width = 0.5)
  ) +
  # Step 3: Add dummy invisible point for legend
  geom_point(
    data = data.frame(Dataset = "dummy", value = 1, Method = "Unconstrained", model = NA),
    aes(x = Dataset, y = value, color = Method, shape = model),
    inherit.aes = FALSE,
    alpha = 0
  ) +
  # Step 4: Set y and color scales
  scale_y_log10(breaks = c(1000, 10, 0.1, 0.001, 1e-5), labels = yLabeler) +
  scale_color_manual(values = c(my_colors, Unconstrained = "black")) +

  # Step 5: Remove "dummy" from x-axis
  scale_x_discrete(limits = dataset_levels) +

  # Step 6: Styling
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.position = "top",
    legend.key = element_rect(fill = NA, colour = NA, linewidth = 0.25),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
scatter_plot

```

```{r}
#load("~/Documents/projects/fixed-tree-experiments/lassaL_kli22.RData")
make_boxplot <- function(dataset,parame,modele, xlabel){
load(paste0(dataset, ".RData"))
df3 = param %>%
  #slice(seq(floor(0.5 * n()), n(), by = 5)) %>%
  select(all_of(parame), type) %>% gather("param", "value", -type)

df3$type <- recode(df3$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="MRC",
                        "unconstrained1"="Unconstrained")

plt = ggplot(df3, aes(value, col = type)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  #geom_density() +
  geom_boxplot(outliers=F) +
  scale_color_manual(values = my_colors) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.position="top",
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    #axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1),
    axis.title.y=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()
  ) +
  labs( x=xlabel) + guides(col="none")

highlight_points = df_params %>% filter(Dataset==dataset& model==modele)

  rectangle = data.frame(
    xmin = which(order_dataset==dataset) - 0.4,
    xmax = which(order_dataset==dataset) + 0.4,
    ymin = min(highlight_points$value),
    ymax = max(highlight_points$value)
  )
  list(plot=plt, rectangle=rectangle)
}

root_lassa = make_boxplot("lassaL_kli22", parame="rootHeight", modele="tree", xlabel="Root age")
rate_wnv = make_boxplot("wnv_del20", parame="meanRate", modele="clock", xlabel="Mean rate")
root_lassa
rate_wnv
```

```{r}
#load("~/Documents/projects/fixed-tree-experiments/lassaL_kli22.RData")
make_boxplot_grid <- function(datasets, params, models, xlabels){
  df3 = NULL
  for(i in seq_along(datasets)){
load(paste0(datasets[i], ".RData"))
df3 = param %>%
  #slice(seq(floor(0.5 * n()), n(), by = 5)) %>%
  select(all_of(params[i]), type) %>% gather("param", "value", -type) %>%
  mutate(dataset=datasets[i], label=xlabels[i]) %>%
  bind_rows(df3)

highlight_points = df_params %>% filter(Dataset==datasets[i]& model==models[i])

  rectangle = data.frame(
    xmin = which(order_dataset==datasets[i]) - 0.4,
    xmax = which(order_dataset==datasets[i]) + 0.4,
    ymin = min(highlight_points$value),
    ymax = max(highlight_points$value)
  )
}
df3$type <- recode(df3$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="MRC",
                        "unconstrained1"="Unconstrained")

plt = ggplot(df3, aes(x=type,y=value, col = type)) +
  geom_boxplot(outliers=F) +
  scale_color_manual(values = my_colors) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.position="top",
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    strip.placement = "outside",
    
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1),
    #axis.title.x=element_blank(),
    #axis.text.x=element_blank(),
    #axis.ticks.x=element_blank()
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank()
  ) +
  facet_wrap(vars(dataset), scale="free", nrow=2, 
                strip.position = "bottom", 
                labeller = as_labeller(c(lassaL_kli22 = "Root age", wnv_del20 = "Mean rate"))) +
  labs( x=NULL,y=NULL) + 
    guides(col="none") + coord_flip()

  plt
}

#root_lassa = make_boxplot("lassaL_kli22", parame="rootHeight", modele="tree", xlabel="Root age")
#rate_wnv = make_boxplot("wnv_del20", parame="meanRate", modele="clock", xlabel="Mean rate")

make_boxplot_grid(c("lassaL_kli22", "wnv_del20"), c("rootHeight", "meanRate"), c("tree","clock"), c("Root age","Mean rate"))
```


```{r}

make_skyplot <- function(dataset1, grid_points, youngest){
  df_temp = data.frame(pop=paste0("skygrid.logPopSize", seq(1,50)), grid=grid_points)
  pop = sky_mean %>% filter(dataset==dataset1) %>% left_join(df_temp, by="pop") %>% mutate(value=exp(value))
  
  pop$type <- recode(pop$type,
                          "mcc" = "MCC",
                          "lsd2" = "LSD",
                          "treetime" = "TreeTime",
                          "map"="MSP",
                          "rerooted"="MRC",
                          "unconstrained1"="Unconstrained")
  
  plt = ggplot(pop, aes(x=grid, y=value, col=type))+
    theme_bw() +
    #scale_y_log10() +
    scale_color_manual(values = my_colors) +
    theme(
      panel.grid = element_blank(),
      axis.line = element_line(linewidth = 0.4),
      strip.background = element_blank(),
      panel.border = element_blank(),
      legend.title = element_blank(),
      legend.background = element_blank(),
      legend.key = element_rect(
        fill = NA,
        colour = NA,
        linewidth = 0.25
      ),
      axis.text.x = element_text(angle = 30, hjust = 1)
    ) +
    geom_line() +
    scale_x_reverse() +
    #scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x), labels = trans_format("log10", math_format(10^.x))) +
    xlab("Year") +
    ylab("Mean effective population size")  + guides(col="none")
  
  highlight_points <- df.pop %>% filter(Dataset==dataset1, model=="coalescent")
  
  rectangle <- data.frame(
    #xmin = min(highlight_points$x),
    #xmax = max(highlight_points$x),
    xmin = which(order_dataset==dataset1) - 0.4,
    xmax = which(order_dataset==dataset1) + 0.4,
    ymin = min(highlight_points$value),
    ymax = max(highlight_points$value)
  )
  
  list(plot=plt, rectangle=rectangle)
}

sky_ebov_mba = make_skyplot('ebov_mba21', rev(2020.1284153005465 - seq(0, 1.75, length.out = 50)))
#sky_ebov_dud = make_skyplot('ebov_dud17', rev(2020.327868852459 - seq(0, 2, length.out = 50)))
sky_pb2_wor = make_skyplot('fluPb2_wor14', rev(2011.4 - seq(0, 150, length.out = 50)))
sky_ebov_mba
sky_pb2_wor
```
```{r}
make_skyplot_grid <- function(){
  df_temp_mba = data.frame(pop=paste0("skygrid.logPopSize", seq(1,50)), grid=rev(2020.1284153005465 - seq(0, 1.75, length.out = 50))) %>% mutate(dataset="ebov_mba21")
  df_temp_wor = data.frame(pop=paste0("skygrid.logPopSize", seq(1,50)), grid=rev(2011.4 - seq(0, 150, length.out = 50))) %>% mutate(dataset="fluPb2_wor14")
  df_temp = rbind(df_temp_mba,df_temp_wor)
  pop = sky_mean %>% filter(dataset%in% c('ebov_mba21', 'fluPb2_wor14')) %>% left_join(df_temp, by=c("pop","dataset")) %>% mutate(value=exp(value))
  
  pop$type <- recode(pop$type,
                          "mcc" = "MCC",
                          "lsd2" = "LSD",
                          "treetime" = "TreeTime",
                          "map"="MSP",
                          "rerooted"="MRC",
                          "unconstrained1"="Unconstrained")
  
  plt = ggplot(pop, aes(x=grid, y=value, col=type))+
    theme_bw() +
    #scale_y_log10() +
    scale_color_manual(values = my_colors) +
    theme(
      panel.grid = element_blank(),
      axis.line = element_line(linewidth = 0.4),
      strip.background = element_blank(),
      panel.border = element_blank(),
      legend.title = element_blank(),
      legend.background = element_blank(),
      legend.key = element_rect(
        fill = NA,
        colour = NA,
        linewidth = 0.25
      ),
      axis.text.x = element_text(angle = 30, hjust = 1)
    ) +
    geom_line() +
    facet_wrap(vars(dataset), scales="free", nrow=2) +
    scale_x_reverse() +
    xlab("Year") +
    ylab("Mean effective population size")  + guides(col="none")
  plt
}

make_skyplot_grid()
```

```{r subpanels}
left = ggarrange(sky_ebov_mba$plot+ylab(NULL), sky_pb2_wor$plot+ylab(NULL),
          labels = c("A", "B"),label.x = 0.3,#common.legend = TRUE, legend = "right"
          ncol = 1, nrow = 2, align = "v") + theme(plot.margin = margin(5, 5, 8, 5))
left = annotate_figure(left, left = text_grob("Mean effective population size", rot = 90))
#left = make_skyplot_grid()
right = ggarrange(root_lassa$plot, rate_wnv$plot,
          labels = c("D", "E"), label.x = 0.05,#-0.07,
          ncol = 1, nrow = 2)
#right = make_boxplot_grid(c("lassaL_kli22", "wnv_del20"), c("rootHeight", "meanRate"), c("tree","clock"), c("Root age","Mean rate"))

scatter_plot2 = scatter_plot + theme(plot.margin = margin(5, 5, 9, 5)) +
  geom_rect(data = root_lassa$rect,
            inherit.aes = FALSE,
            aes(xmin = xmin+0.3, xmax = xmax-0.3, ymin = ymin*0.5, ymax = ymax+5),
            fill = NA, color = "black", linewidth = 0.5)+
  annotate(
    "text", label = "D",
    x = root_lassa$rect$xmax-0.3, y = root_lassa$rect$ymax*3, size = 4, colour = "black"
  ) +
  geom_rect(data = rate_wnv$rect, inherit.aes = FALSE,
            aes(xmin = xmin+0.1, xmax = xmax-0.47, ymin = ymin*0.5, ymax = ymax*1.5),
            fill = NA, color = "black", linewidth = 0.5)+
  annotate(
    "text", label = "E",
    x = rate_wnv$rect$xmax-0.6, y = rate_wnv$rect$ymax*3, size = 4, colour = "black"
  ) +
  geom_rect(data = sky_ebov_mba$rect,
            inherit.aes = FALSE,
            aes(xmin = xmin+0.45, xmax = xmax-0.1, ymin = ymin*0.5, ymax = ymax*1.5),
            fill = NA, color = "black", linewidth = 0.5) +
  annotate(
    "text", label = "A",
    x = sky_ebov_mba$rect$xmax, y = sky_ebov_mba$rect$ymax*3, size = 4, colour = "black"
  ) +
  geom_rect(data = sky_pb2_wor$rect, inherit.aes = FALSE,
            aes(xmin = xmin+0.45, xmax = xmax-0.1, ymin = ymin*0.05, ymax = ymax*1.5),
            fill = NA, color = "black", linewidth = 0.5) +
  annotate(
    "text", label = "B",
    x = sky_pb2_wor$rect$xmax, y = sky_pb2_wor$rect$ymax*3, size = 4, colour = "black"
  )
  
figure = ggarrange(left, scatter_plot2, right,
          labels = c("", "C"), label.x = 0.07,
          widths = c(1,2,1),
          ncol = 3, nrow = 1,  common.legend = TRUE)
figure
#annotate_figure(figure, left = text_grob("Mean effective population size", rot = 90))
```

