---
title: "fixed tree: summariy"
author: "Mathieu"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(scales)
```

\newpage

```{r}
kld_base = function(x, y, ...) {
  integrand = function(x, y, t) {
    f.x =  approx(density(x)$x, density(x)$y, t)$y
    f.y =  approx(density(y)$x, density(y)$y, t)$y
    tmpRatio = f.x * (log2(f.x) - log2(f.y))
    tmpRatio = ifelse(is.infinite(tmpRatio), 0, ifelse(is.na(tmpRatio), 0, tmpRatio))
    return(tmpRatio)
  }
  return(integrate(
    integrand,
    0,
    Inf,
    x = x,
    y = y,
    stop.on.error = FALSE
  )$value)
}

kl_l <- function(x, y) {
  f.x <- approxfun(density(x, bw = "nrd0"))
  f.y <- approxfun(density(y, bw = "nrd0"))
  integrand <- function(z) {
    return((log(f.x(z)) - log(f.y(z))) * f.x(z))
  }
  return(integrate(integrand, lower = 2, upper = 100))
}
```

```{r eval=FALSE, include=FALSE}
# ebov_dud17
#c("cp1.kappa", "ig.kappa", "cp1.alpha", "ig.alpha", "cp1.mu", "ig.mu", "cp1.frequencies","ig.frequencies")
#c("ucld.mean", "ucld.stdev", "meanRate")

# ebov_mba21
# c("cp12.kappa", "cp3.kappa", "cp12.alpha", "cp3.alpha", "cp12.mu", "cp3.mu", "cp12.frequencies", "cp3.frequencies")
# c("ucld.mean", "ucld.stdev", "meanRate")

# fluH1L_bed15
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c("constant.popSize", "clock.rate")

# fluH3L_bed15
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c("constant.popSize", "clock.rate")

# fluPb2_wor14
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c(equine.rate","galliform1.rate","human1.rate","shorebird1.rate","swine1.rate","swine2.rate","swine3.rate")

# fluVicL_bed15
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c("ucld.mean", "ucld.stdev", "meanRate")

# hiv_far14
# c("gtr.rates", "frequencies", "alpha")
# c("ucld.mean", "ucld.stdev", "meanRate")

# lassaL_kli22
# c("gtr.rates", "frequencies", "alpha")
# c("ucld.mean", "ucld.stdev", "meanRate")

# mumps_mon21
# c("gtr.rates", "frequencies", "alpha")
# c("clock.rate")

# rabies_via23
# c("cp12.gtr.rates", "cp3.gtr.rates", "nc.gtr.rates", "cp12.alpha", "cp3.alpha", "nc.alpha", "cp12.frequencies" "cp3.frequencies", "nc.frequencies", "cp12.mu", "cp3.mu", "nc.mu")
# c("ucld.mean", "ucld.stdev", "meanRate")

# sars2_can20
# c("kappa", "frequencies", "alpha")
# c("clock.rate")

# sars2_lem21
# c("kappa", "frequencies", "alpha")
# c("clock.rate")

# sars2_pek22
# c("gtr.rates", "frequencies", "pInv")
# c("clock.rate")

# wnv_del20
# c("gtr.rates", "frequencies", "alpha")
# c("ucld.mean", "ucld.stdev", "meanRate")

# zikv_gru19
# c("cp1.kappa", "cp2.kappa", "cp3.kappa", "cp1.alpha", "cp2.alpha", "cp3.alpha", "cp1.mu", "cp2.mu", "cp3.mu", "cp1.frequencies", "cp2.frequencies", "cp3.frequencies"/>
# c("ucld.mean", "ucld.stdev", "meanRate")
```

```{r}
datasets = c("fluH3L_bed15", "lassaL_kli22", "wnv_del20", "sars2_lem21", "ebov_dud17",
  "ebov_mba21", "fluH1L_bed15", "fluPb2_wor14", "fluVicL_bed15", "hiv_far14",
  "mumps_mon21", "rabies_via23", "sars2_can20", "sars2_pek22", "zikv_gru19")
all = NULL
all_mean = NULL
all_sky = NULL
for (i in seq_along(datasets)) {
  dataset = datasets[[i]]
  load(paste0(dataset, ".RData"))
  rate_name <- intersect(c("meanRate", "clock.rate", "avian.rate"), names(param))
  param = param %>% mutate(!!"rate" := .[[rate_name[1]]])
  
  df = param %>% group_by(type) %>%
    summarize(
      rootHeightMean = mean(rootHeight),
      rootHeightVar = var(rootHeight),
      rateMean = mean(rate)
    ) %>%
    mutate(dataset = dataset)

  all_mean = param %>% group_by(type) %>%
    summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
    pivot_longer(!type, names_to = "parameter", values_to = "value") %>%
    mutate(dataset = dataset) %>%
    bind_rows(all_mean)
  
  all = rbind(all, df)
  #assign(paste0("param_", dataset), get("param"))
  #assign(paste0("sky_", dataset), get("sky"))
  #assign(paste0("lsd_", dataset), get("lsd"))
  if (!is.null(sky)) {
    prefixes <- names(sky)[1:length(sky) - 1]
    m = sky %>% group_by(type) %>%
      summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)))
    uncons = m %>% filter(type == "unconstrained1") %>%
      pivot_longer(cols = all_of(prefixes),
                   names_to = "pop",
                   values_to = "value")
    cons = m %>% filter(type != "unconstrained1") %>%
      pivot_longer(cols = all_of(prefixes),
                   names_to = "pop",
                   values_to = "value")
    
    df = cons %>% full_join(uncons, by = "pop") %>%
      mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
      select(type.x, value) %>%
      # RRMSE relative root mean squared error
      group_by(type.x) %>% summarize(across(where(is.numeric), ~ sqrt(mean(.)))) %>%
      rename(Method = type.x, RRMSE = value) %>%
      select(Method, RRMSE) %>%
      mutate(Dataset = dataset)
    
    all_sky = df %>% bind_rows(all_sky)
  }
}
```

```{r}
uncons = all_mean %>% filter(type == "unconstrained1") %>% select(-type)
cons = all_mean %>% filter(type != "unconstrained1")
df = cons %>% full_join(uncons, by = c("dataset", "parameter")) %>%
  mutate(value = (value.x - value.y) / value.y) %>%
  select(type, parameter, dataset, value) %>%
  rename(Method = type, Dataset = dataset) %>%
  filter(parameter != "rate") %>%
  mutate(
    model = case_when(
      parameter == "meanRate" |
        parameter == "clock.rate" | str_detect(parameter, ".rate$") ~ "clock",
      parameter == "rootHeight" |
        parameter == "treeLength" | str_detect(parameter, "^tmrca") ~ "tree",
      parameter == "constant.popSize" |
        parameter == "skygrid.precision" ~ "coalescent",
      str_detect(parameter, "gtr") |
        str_detect(parameter, "mu") |
        str_detect(parameter, "frequencies") |
        str_detect(parameter, "kappa") ~ "subsmodel",
      str_detect(parameter, "mu") |
        str_detect(parameter, "alpha") | parameter == "pInv" ~ "sitemodel",
      TRUE                            ~ "other"
    )
  )
```

# Substitution and site model parameters

- Substitution model: kappa, gtr.rates, frequencies
- Site model: alpha, pInv, mu

```{r}
df2 = df %>% filter(model %in% c("other", "subsmodel", "sitemodel"))
ggplot(df2, aes(y = value, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  theme(
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

# Clock, coalescent, tree parameters/statistics

- Coalescent: constant popSize, precision of GMRF
- Tree: rootHeight, treeLength, tmrcas
- Clock model: strict clock rate, meanRate (UCLN), *.rate of local clock

```{r}
df2 = df %>% filter(model %in% c("coalescent", "tree", "clock"))
ggplot(df2, aes(y = value, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  theme(
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

# Skygrid population size parameters

$RRMSE(\bf{\bar x}) = \sqrt{1/n \sum_i^n \left(\frac{\bar x_i - \bar x_i^{u}}{\bar x_i^{u}}\right)^2}$ where $\bar x_i^{u}$ is the $i^{th}$ mean population size parameter of the unconstrained run and $\bar x_i$ is the mean population size parameter for a constrained run.

```{r}
ggplot(all_sky, aes(y = RRMSE, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1) +
  theme(
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```
