---
title: "fixed tree: summarry"
author: "Mathieu"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)
knitr::opts_chunk$set(fig.path='Figs/')
```

\newpage

```{r}
kld_base = function(x, y, ...) {
  integrand = function(x, y, t) {
    f.x =  approx(density(x)$x, density(x)$y, t)$y
    f.y =  approx(density(y)$x, density(y)$y, t)$y
    tmpRatio = f.x * (log2(f.x) - log2(f.y))
    tmpRatio = ifelse(is.infinite(tmpRatio), 0, ifelse(is.na(tmpRatio), 0, tmpRatio))
    return(tmpRatio)
  }
  return(integrate(
    integrand,
    0,
    Inf,
    x = x,
    y = y,
    stop.on.error = FALSE
  )$value)
}

kl_l <- function(x, y) {
  f.x <- approxfun(density(x, bw = "nrd0"))
  f.y <- approxfun(density(y, bw = "nrd0"))
  integrand <- function(z) {
    return((log(f.x(z)) - log(f.y(z))) * f.x(z))
  }
  return(integrate(integrand, lower = 2, upper = 100))
}
```

```{r eval=FALSE, include=FALSE}
# ebov_dud17
#c("cp1.kappa", "ig.kappa", "cp1.alpha", "ig.alpha", "cp1.mu", "ig.mu", "cp1.frequencies","ig.frequencies")
#c("ucld.mean", "ucld.stdev", "meanRate")

# ebov_mba21
# c("cp12.kappa", "cp3.kappa", "cp12.alpha", "cp3.alpha", "cp12.mu", "cp3.mu", "cp12.frequencies", "cp3.frequencies")
# c("ucld.mean", "ucld.stdev", "meanRate")

# fluH1L_bed15
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c("constant.popSize", "clock.rate")

# fluH3L_bed15
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c("constant.popSize", "clock.rate")

# fluPb2_wor14
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c(equine.rate","galliform1.rate","human1.rate","shorebird1.rate","swine1.rate","swine2.rate","swine3.rate")

# fluVicL_bed15
# c("cp12.kappa", "cp3.kappa", "cp12.frequencies", "cp3.frequencies", "cp12.mu", "cp3.mu")
# c("ucld.mean", "ucld.stdev", "meanRate")

# hiv_far14
# c("gtr.rates", "frequencies", "alpha")
# c("ucld.mean", "ucld.stdev", "meanRate")

# lassaL_kli22
# c("gtr.rates", "frequencies", "alpha")
# c("ucld.mean", "ucld.stdev", "meanRate")

# mumps_mon21
# c("gtr.rates", "frequencies", "alpha")
# c("clock.rate")

# rabies_via23
# c("cp12.gtr.rates", "cp3.gtr.rates", "nc.gtr.rates", "cp12.alpha", "cp3.alpha", "nc.alpha", "cp12.frequencies" "cp3.frequencies", "nc.frequencies", "cp12.mu", "cp3.mu", "nc.mu")
# c("ucld.mean", "ucld.stdev", "meanRate")

# sars2_can20
# c("kappa", "frequencies", "alpha")
# c("clock.rate")

# sars2_lem21
# c("kappa", "frequencies", "alpha")
# c("clock.rate")

# sars2_pek22
# c("gtr.rates", "frequencies", "pInv")
# c("clock.rate")

# wnv_del20
# c("gtr.rates", "frequencies", "alpha")
# c("ucld.mean", "ucld.stdev", "meanRate")

# zikv_gru19
# c("cp1.kappa", "cp2.kappa", "cp3.kappa", "cp1.alpha", "cp2.alpha", "cp3.alpha", "cp1.mu", "cp2.mu", "cp3.mu", "cp1.frequencies", "cp2.frequencies", "cp3.frequencies"/>
# c("ucld.mean", "ucld.stdev", "meanRate")
```

```{r eval=FALSE, include=FALSE}
datasets = c("fluH3L_bed15", "lassaL_kli22", "wnv_del20", "sars2_lem21", "ebov_dud17",
  "ebov_mba21", "fluH1L_bed15", "fluPb2_wor14", "fluVicL_bed15", "hiv_far14",
  "mumps_mon21", "rabies_via23", "sars2_can20", "sars2_pek22", "zikv_gru19")
all = NULL
all_mean = NULL
all_sky = NULL
for (i in seq_along(datasets)) {
  dataset = datasets[[i]]
  load(paste0(dataset, ".RData"))
  rate_name <- intersect(c("meanRate", "clock.rate", "avian.rate"), names(param))
  param = param %>% mutate(!!"rate" := .[[rate_name[1]]]) %>% filter(type !="fixed")
  
  df = param %>% group_by(type) %>%
    summarize(
      rootHeightMean = mean(rootHeight),
      rootHeightVar = var(rootHeight),
      rateMean = mean(rate)
    ) %>%
    mutate(dataset = dataset)

  all_mean = param %>% group_by(type) %>%
    summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
    pivot_longer(!type, names_to = "parameter", values_to = "value") %>%
    mutate(dataset = dataset) %>%
    bind_rows(all_mean)
  
  all = rbind(all, df)
  #assign(paste0("param_", dataset), get("param"))
  #assign(paste0("sky_", dataset), get("sky"))
  #assign(paste0("lsd_", dataset), get("lsd"))
  if (!is.null(sky)) {
    prefixes <- names(sky)[1:length(sky) - 1]
    m = sky %>% group_by(type) %>%
      summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)))
    uncons = m %>% filter(type == "unconstrained1") %>%
      pivot_longer(cols = all_of(prefixes),
                   names_to = "pop",
                   values_to = "value")
    cons = m %>% filter(type != "unconstrained1") %>%
      pivot_longer(cols = all_of(prefixes),
                   names_to = "pop",
                   values_to = "value")
    
    df = cons %>% full_join(uncons, by = "pop") %>%
      mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
      select(type.x, value) %>%
      # RRMSE relative root mean squared error
      group_by(type.x) %>% summarize(across(where(is.numeric), ~ sqrt(mean(.)))) %>%
      rename(Method = type.x, RRMSE = value) %>%
      select(Method, RRMSE) %>%
      mutate(Dataset = dataset)
    
    all_sky = df %>% bind_rows(all_sky)
  }
}
```

```{r eval=FALSE, include=FALSE}
datasets = c(
  "fluH3L_bed15",
  "lassaL_kli22",
  "wnv_del20",
  "sars2_lem21",
  "ebov_dud17",
  "ebov_mba21",
  "fluH1L_bed15",
  "fluPb2_wor14",
  "fluVicL_bed15",
  "hiv_far14",
  "mumps_mon21",
  "rabies_via23",
  "sars2_can20",
  "sars2_pek22",
  "zikv_gru19"
)
#c("map"="HPT", "lsd2"="LSD", "mcc"= "MCC", "treetime"="TreeTime", "unconstrained"="Unconstrained", "rerooted"="IQ-TREE/rooted")
all_mean = NULL
all_sd = NULL
sky_mean = NULL
sky_sd = NULL

for (dataset in datasets) {
  for (topology in c("unconstrained", "lsd2", "treetime", "map", "mcc", "rerooted")) {
    if (topology == "unconstrained") {
      stem = file.path("ssstree_datasets", dataset)
      log.file = file.path(stem, paste0("param_run0", 1, ".log"))
    } else{
      stem = file.path("results", "datasets", dataset, topology)
      log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      if (!file.exists(log.file)) {
        stem = file.path("results-lsd", "datasets", dataset, topology)
        log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      }
      if (!file.exists(log.file)) {
        stem = file.path("results-treetime", "datasets", dataset, topology)
        log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      }
      if (!file.exists(log.file)) {
        stem = file.path("results-treetime", "datasets", dataset)
        log.file = file.path(stem, paste0("param_run", "0", 1, ".log"))
      }
    }
    if (!file.exists(log.file)) {
      next
    }
    
    df.log <- read_tsv(log.file, skip = 3, show_col_types = FALSE)
    df.log = df.log %>%
      slice(seq(floor(0.5 * n()), n(), by=5)) %>%
      select(-state)
    #df.log <- df.log[seq(1, nrow(df.log), 5), ] # subsample
    #df.log <- df.log[as.integer(0.5 * nrow(df.log)):nrow(df.log), ] %>% select(-state)
    
    all_mean = df.log %>%
      summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
      pivot_longer(cols = everything(),
                   names_to = "parameter",
                   values_to = "value") %>%
      mutate(dataset = dataset, type = topology) %>%
      bind_rows(all_mean)
    
    all_sd = df.log %>%
      summarise(across(where(is.numeric), ~ sd(.x, na.rm = TRUE))) %>%
      pivot_longer(cols = everything(),
                   names_to = "parameter",
                   values_to = "value") %>%
      mutate(dataset = dataset, type = topology) %>%
      bind_rows(all_sd)
    
    if (file.exists(file.path(stem, paste0("sky_run", "0", 1, ".log")))) {
      sky.log <- read_tsv(file.path(stem, paste0("sky_run", "0", 1, ".log")),
                          skip = 3,
                          show_col_types = FALSE)
      sky.log = sky.log %>%
        slice(seq(floor(0.5 * n()), n(), by=5)) %>%
        select(-state)

      sky_mean = sky.log %>%
        summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE))) %>%
        pivot_longer(
          cols = everything(),
          names_to = "pop",
          values_to = "value"
        ) %>%
        mutate(dataset = dataset, type = topology) %>%
        bind_rows(sky_mean)
      
      sky_sd = sky.log %>%
        summarise(across(where(is.numeric), \(x) sd(x, na.rm = TRUE))) %>%
        pivot_longer(
          cols = everything(),
          names_to = "pop",
          values_to = "value"
        ) %>%
        mutate(dataset = dataset, type = topology) %>%
        bind_rows(sky_sd)
    }
  }
}
save(all_mean, all_sd, sky_mean, sky_sd, file = "summary.RData")
```

```{r eval=FALSE, include=FALSE}
for(dataset in c("rabies_via23")){
  for(type in c()){
df.log <- read_tsv(file.path("results-treetime", "datasets", dataset, paste0("param_run", "0", 1, ".log")),
    skip = 3, show_col_types = FALSE
  )
rate_name <- intersect(c("meanRate", "clock.rate", "avian.rate"), names(df.log))
  df.log <- df.log[seq(1, nrow(df.log), 5), ]
  param2 <- df.log[as.integer(0.5 * nrow(df.log)):nrow(df.log), ] %>%
    #select(any_of(c("rootHeight", "clock.rate", "treeLength", "pInv", "skygrid.precision"))) %>% # sars2_pek22
    select(any_of(c("rootHeight", "meanRate", "treeLength", "cp12.alpha", "cp12.mu", "skygrid.precision"))) %>% #rabies_via23
    mutate(type = "rerooted") %>% mutate(!!"rate" := .[[rate_name[1]]])
  
  all_mean = param2 %>% group_by(type) %>%
    summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>%
    pivot_longer(!type, names_to = "parameter", values_to = "value") %>%
    mutate(dataset = dataset) %>%
    bind_rows(all_mean)
  }
}
```

```{r}
load("summary.RData")
all_mean$type <- recode(all_mean$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="Rerooted")

all_sd$type <- recode(all_sd$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="Rerooted")

sky_mean$type <- recode(sky_mean$type,
                        "mcc" = "MCC",
                        "lsd2" = "LSD",
                        "treetime" = "TreeTime",
                        "map"="MSP",
                        "rerooted"="Rerooted")
```

```{r}
all_mean = all_mean %>% mutate(stat="mean")
all_sd = all_sd %>% mutate(stat="sd")

uncons = all_mean %>% bind_rows(all_sd) %>% filter(type == "unconstrained") %>% select(-type)
cons = all_mean %>% bind_rows(all_sd) %>% filter(type != "unconstrained")

df_all = cons %>% full_join(uncons, by = c("dataset", "parameter", "stat")) %>%
  mutate(value = (value.x - value.y) / value.y) %>%
  select(type, parameter, dataset, value, stat) %>%
  rename(Method = type, Dataset = dataset) %>%
  filter(parameter != "rate") %>%
  mutate(
    model = case_when(
      parameter == "meanRate" |
        parameter == "clock.rate" | str_detect(parameter, ".rate$") ~ "clock",
      parameter == "rootHeight" |
        parameter == "treeLength" | str_detect(parameter, "^tmrca") ~ "tree",
      parameter == "constant.popSize" |
        parameter == "skygrid.precision" ~ "coalescent",
      str_detect(parameter, "gtr") |
        str_detect(parameter, "mu") |
        str_detect(parameter, "frequencies") |
        str_detect(parameter, "kappa") ~ "subsmodel",
      str_detect(parameter, "mu") |
        str_detect(parameter, "alpha") | parameter == "pInv" ~ "sitemodel",
      TRUE                            ~ "other"
    )
  )
df_mean = df_all %>% filter(stat == "mean")
df_sd = df_all %>% filter(stat == "sd")
```

```{r}
rerooted = df_mean %>% filter(Method == "Rerooted")
same = df_mean %>% filter(Method != "Rerooted") %>% filter(Method %in% c("LSD", "TreeTime"))

df2=rerooted %>% left_join(same, by = c("Dataset", "parameter", "stat", "model")) %>%
  rename(rerooted=value.x, heuristic=value.y) %>%
  mutate(diff=rerooted-heuristic) %>%
  filter(model!= "other")

ggplot(df2, aes(rerooted, heuristic, col=model, shape=Method.y))+
  geom_point() +
  geom_abline(slope=1,intercept=0) #+
  #facet_wrap(~model)
```
```{r}
print(paste0(nrow(filter(df2, Method.y=="LSD" & diff < 0)), " ", nrow(filter(df2, Method.y=="LSD"))))
print(paste0(nrow(filter(df2, Method.y=="TreeTime" & diff < 0)), " ", nrow(filter(df2, Method.y=="TreeTime"))))

print(paste0(nrow(filter(df2, Method.y=="LSD" & diff < 0 & model %in% c("coalescent", "tree", "clock"))), " ", nrow(filter(df2, Method.y=="LSD",model %in% c("coalescent", "tree", "clock")))))
print(paste0(nrow(filter(df2, Method.y=="TreeTime" & diff < 0 & model %in% c("coalescent", "tree", "clock"))), " ", nrow(filter(df2, Method.y=="TreeTime",model %in% c("coalescent", "tree", "clock")))))
```

```{r}
df2=cons %>% full_join(uncons, by = c("dataset", "parameter", "stat")) %>%
  rename(constrained=value.x, unconstrained=value.y) %>%
  filter(parameter!="state"& parameter!="treeLength") %>%
  mutate(
    model = case_when(
      parameter == "meanRate" |
        parameter == "clock.rate" | str_detect(parameter, ".rate$") ~ "clock",
      parameter == "rootHeight" |
        parameter == "treeLength" | str_detect(parameter, "^tmrca") ~ "tree",
      parameter == "constant.popSize" |
        parameter == "skygrid.precision" ~ "coalescent",
      str_detect(parameter, "gtr") |
        str_detect(parameter, "mu") |
        str_detect(parameter, "frequencies") |
        str_detect(parameter, "kappa") ~ "subsmodel",
      str_detect(parameter, "mu") |
        str_detect(parameter, "alpha") | parameter == "pInv" ~ "sitemodel",
      TRUE                            ~ "other"
    )
  ) %>%
  filter(stat=="mean")

ggplot(df2, aes(x=unconstrained, y=constrained, col=type, shape=model)) +
  geom_point() +
  geom_abline(slope=1,intercept=0)
```

# Substitution and site model parameters

- Substitution model: kappa, gtr.rates, frequencies
- Site model: alpha, pInv, mu

```{r insensitive}
df2 = df_mean %>% filter(model %in% c("subsmodel", "sitemodel"))
#pdf("Figs/insensitive.pdf")
ggplot(df2, aes(y = value, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "Relative error")
#dev.off()
```

# Clock, coalescent, tree parameters/statistics

- Coalescent: constant popSize, precision of GMRF
- Tree: rootHeight, treeLength, tmrcas
- Clock model: strict clock rate, meanRate (UCLN), *.rate of local clock

```{r sensitive}
df2 = df_mean %>% filter(model %in% c("coalescent", "tree", "clock"))  %>%
  filter(!parameter %in% c("meanRate", "treeLength"))
#pdf("Figs/sensitive.pdf")
ggplot(df2, aes(y = value, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1, aes(shape = model)) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "Relative error")
```

```{r re-mean}
df2 = df_mean %>%
  filter(model %in% c("coalescent", "tree", "clock", "subsmodel", "sitemodel")) %>%
  filter(!parameter %in% c("meanRate", "treeLength"))

#pdf("Figs/sensitive.pdf")
ggplot(df2, aes(y = value, x = model, col = Method)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = Dataset)) +
  geom_boxplot() +
  #geom_jitter(size = 0.5) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(0.1, 0.8),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank()
  ) +
  labs(y = "Relative error of mean") #+ ggtitle("Relative error of mean")
#dev.off()
```

```{r}
df2 = df_sd %>% filter(model %in% c("coalescent", "tree", "clock", "subsmodel", "sitemodel")) %>%
  filter(!parameter %in% c("meanRate", "treeLength"))

#pdf("Figs/sensitive.pdf")
ggplot(df2, aes(y = value, col = Method, x = model)) +
  #geom_jitter(width = 0.2, size = 1, aes(shape = Dataset)) +
  geom_boxplot() +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "Relative error") +
  ggtitle("Relative error of standard deviation")
#dev.off()
```

# Skygrid population size parameters

$RRMSE(\bf{\bar x}) = \sqrt{1/n \sum_i^n \left(\frac{\bar x_i - \bar x_i^{u}}{\bar x_i^{u}}\right)^2}$ where $\bar x_i^{u}$ is the $i^{th}$ mean population size parameter of the unconstrained run and $\bar x_i$ is the mean population size parameter for a constrained run.

```{r skygrids}
#pdf("Figs/skygrids.pdf")
uncons = sky_mean %>% filter(type == "unconstrained")
cons = sky_mean %>% filter(type != "unconstrained")

df.pop = cons %>% full_join(uncons, by = c("dataset", "pop")) %>%
  mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
  select(type.x, value, dataset) %>%
  group_by(type.x, dataset) %>% summarise(sqrt_mean = sqrt(mean(value)), .groups = "drop") %>%
  rename(Method = type.x, RMSRE = sqrt_mean, Dataset = dataset)# %>% filter(RRMSE <100)


ggplot(df.pop, aes(y = RMSRE, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1) +
  scale_y_log10() +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
#dev.off() 
```

```{r}
#pdf("Figs/skygrids.pdf")
uncons = sky_mean %>% filter(type == "unconstrained")
cons = sky_mean %>% filter(type != "unconstrained")

df.pop = cons %>% full_join(uncons, by = c("dataset", "pop")) %>%
  mutate(value = ((value.x - value.y) / value.y) ** 2) %>%
  select(type.x, value, dataset) %>%
  group_by(type.x, dataset) %>% summarise(sqrt_mean = sqrt(mean(value)), .groups = "drop") %>%
  rename(Method = type.x, RRMSE = sqrt_mean, Dataset = dataset) %>% filter(Dataset != "ebov_mba21")


ggplot(df.pop, aes(y = RRMSE, col = Method, x = Dataset)) +
  geom_jitter(width = 0.2, size = 1) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 0.4),
    strip.background = element_blank(),
    panel.border = element_blank(),
    legend.title = element_blank(),
    legend.background = element_blank(),
    legend.key = element_rect(
      fill = NA,
      colour = NA,
      linewidth = 0.25
    ),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1)
  ) +
  labs(y = "RMSRE") + 
  ggtitle("No ebov_mba21")
#dev.off() 
```
